---
title: "ontarioparkscrape"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is set up across multiple pages:


```{r cars}
library(rvest)
library(tidyverse)
library(sf)
library(patchwork)
library(ggstream)
library(ggtext)
```

```{r}
sysfonts::font_add_google(name = "Lato", "Lato")
showtext::showtext_auto()
```

```{r}
#links
baseurl<-"https://en.wikipedia.org/wiki/List_of_provincial_parks_of_"

SWO<-"Southwestern_Ontario"
NO<-"Northern_Ontario"
CO<-"Central_Ontario"
EO<-"Eastern_Ontario"
GH<-"the_Golden_Horseshoe"
```


Change the index of each page and go through each page. This is because there are slight quirks for each page and the table headings differ
```{r}
#sequence of url strings
urlend<-c(SWO,NO,CO,EO,GH)#short name used for assigning
fullurl<-paste0(baseurl,urlend)#full name for webscraping

index<-5
```

Read all tables and headers 

Note for eastern ontario and Golden horseshoe, there is an error becuase these pages had no table of contents. Uncommenting will assign the correct table of contents name
```{r}
#read HTML content from url
content <- read_html(fullurl[index])

#read all tables on the page, and parse into dataframes
tables<- content %>% html_nodes("table") %>% html_table(fill = TRUE)

#remove tables that contain don't tables we want.
#keep only ones that have 'name' as a column
nameslist<-lapply(tables,names)
tableswithoutdata<-which(grepl("Name",nameslist))
tables<-tables[tableswithoutdata]


#get the inner text from the table of contents 
#select divs with class 'toc'.
#within this, select ul tag
#each ul has a nested a tag, select this
#get the inner text
toc<-content %>% html_nodes("div#toc") %>% html_nodes("ul") %>% html_nodes("a") %>% html_text()
#

#eastern ontario page missing table of contents (single table)
#toc<-"Eastern Ontario"
#toc<-"Golden Horseshoe"

if(length(tables)==length(toc)){
  print("tables and TOC are same length")
} else {
  print("tables and toc are not the same length")
}

```

#may be an extra table needing to remove
lapply(tables,names)

######select only columns of interest from all dataframes. These are
######Name, Established, and Coordinates
#using lapply
```{r}
tables_clean<-lapply(tables, function(y) y %>% select(Name,Established,Coordinates))
```

#create new columns, once specifygin the region (using the table of contents lst),
#which is different for each table, and 1 and page url
#I used the purr map2 function to map together 2 lists
```{r}
tabs_w_region<-purrr::map2(tables_clean, toc, ~ mutate(.x, region = .y, url = fullurl[index]))
```


#
#I got an error on second page so had to correct a value manually and convert to integer to do the list to dataframe conversion.

#make sure 'Established' column is all as integer. One had fuller date format which was turned to character
#tabs_w_region[[6]]$Established[2]<-1893
tabs_w_region[[6]]$Established<-
tabs_w_region[[6]]$Established %>% as.integer()


```{r}
#flatten the list of dataframes into one dataframe
flattened<-tabs_w_region %>% bind_rows()
#get decimal lat/long, and split up lat/long
flattened$decimal<-str_split_fixed(flattened$Coordinates, pattern = " / ", n = Inf)[,3]
flattened$lat<-as.numeric(str_split_fixed(flattened$decimal, pattern = ";", n = Inf)[,1])
flattened$long<-as.numeric(str_split_fixed(flattened$decimal, pattern = ";", n = Inf)[,2])
print("single table created")
```

```{r}
assign(urlend[index],flattened)
print(paste0(urlend[index], " df created"))
```

#now go back an do cells with other pages, and combine all
```{r}
alldfs<-bind_rows(Northern_Ontario,
               Southwestern_Ontario,
               Central_Ontario,
               Eastern_Ontario,
               the_Golden_Horseshoe, .id = "url_region")
```

write.csv(alldfs,"C:/Users/jeffz/Desktop/data analysis projects/ont_parks scrape/alldata.csv")


Now that we have the lat/long and dates we will add the 
```{r}
# Add the map -------------------------------------------------------------
canada_cd <- st_read("C:/Users/jeffz/Desktop/data analysis projects/ontarioparks/canada_cd_sim.geojson", quiet = TRUE) # 1
canada_cd$PRNAME<-as.factor(canada_cd$PRNAME)

#crs_string = "+proj=lcc +lat_1=49 +lat_2=77 #+lon_0=-91.52 +x_0=0 +y_0=0 +datum=NAD83 +units=m #+no_defs" # 2

#transform ontariomap

```


Make ontario map and the long/lat the same coordinate/projection system
```{r}

alldfs<-alldfs %>% mutate(yearlab_long = -93,yearlab_lat = 46)

mypoints<-alldfs %>% st_as_sf(coords = c("long", "lat"), crs = 4326) %>% 
  st_transform(crs = 4326) 
  
addlabel<-alldfs %>% st_as_sf(coords = c("yearlab_long", "yearlab_lat"), crs = 4326) %>% 
  st_transform(crs = 4326) 

ontario<- canada_cd[canada_cd$PRNAME=="Ontario",] %>% st_transform(crs = 4326) 

mypoints<-na.omit(mypoints)

mypoints$

```

Clean up region names
```{r}
alldfs$regionclean<-NA
alldfs$regionclean<-alldfs$url
alldfs$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Northern_Ontario","Northern Ontario", alldfs$regionclean)
alldfs$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Southwestern_Ontario","Southwestern Ontario", alldfs$regionclean)
alldfs$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Central_Ontario","Central Ontario", alldfs$regionclean)
alldfs$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Eastern_Ontario","Eastern Ontario", alldfs$regionclean)
alldfs$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_the_Golden_Horseshoe","Golden Horseshoe", alldfs$regionclean)

```

Clean up region names
```{r}
mypoints$regionclean<-NA
mypoints$regionclean<-mypoints$url
mypoints$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Northern_Ontario","Northern Ontario", mypoints$regionclean)
mypoints$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Southwestern_Ontario","Southwestern Ontario", mypoints$regionclean)
mypoints$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Central_Ontario","Central Ontario", mypoints$regionclean)
mypoints$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_Eastern_Ontario","Eastern Ontario", mypoints$regionclean)
mypoints$regionclean<-gsub("https://en.wikipedia.org/wiki/List_of_provincial_parks_of_the_Golden_Horseshoe","Golden Horseshoe", mypoints$regionclean)

```

regions as factors
```{r}
alldfs$regionclean<-factor(alldfs$regionclean, levels = c("Northern Ontario",
                                "Central Ontario",
                                "Southwestern Ontario",
                                "Eastern Ontario",
                                "Golden Horseshoe"))

mypoints$regionclean<-factor(mypoints$regionclean, levels = c("Northern Ontario",
                                "Central Ontario",
                                "Southwestern Ontario",
                                "Eastern Ontario",
                                "Golden Horseshoe"))                               
```
plot ontario map
```{r}
ontariomap<-ggplot(data = mypoints) +
  geom_sf(aes(fill = regionclean), size = 0.1, 
          data = ontario,
          fill = alpha("#0075bf",0.25),
          col = "black") + 
  labs(x = "", y = "")+
  #theme_bw()+
  #geom_text(aes(x = -90, y = 46, aes(label = Established)))+
    geom_sf(data = mypoints, aes(fill = regionclean), shape = 21)+#fill = "#00ab67"
  theme(axis.text = element_blank(),
        panel.border = element_rect(colour = "white", fill=NA, size = 5),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank(),
        plot.background = element_rect(fill = "white", colour = NA))+
annotate("text",x = -90, y =45, label = "")  


ontariomap

#ggplot(data = world) +
#    geom_sf() +
#    geom_sf(data = sites, size = 4, shape = 23, fill = "darkred") +
#    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)

```

plot counts per region
```{r}
regioncounts<-alldfs %>% group_by(regionclean) %>% count() %>% arrange(desc(n)) 

regioncounts$regionclean<-factor(regioncounts$regionclean, levels = c("Northern Ontario",
                                "Central Ontario",
                                "Southwestern Ontario",
                                "Eastern Ontario",
                                "Golden Horseshoe"))

countlabel<-"Most provincial parks are found in Northern Ontario. Few are found in the Golden Horseshoe."

count<-ggplot(regioncounts, aes(fill=regionclean, y=n, x=regionclean)) + 
    geom_bar(stat="identity", width = 0.2, color = "black")+
    geom_text(aes(y = n+7,label = n), color = "black")+
    #coord_polar(theta = "y",start = , clip = "off")+
    theme_void()+
  xlab("Region")+
    theme(legend.position = "none",#c(0.8,0.8),
          legend.title = element_blank(),
    axis.text.x = element_text(color = "black"),
     axis.title.x = element_text(color = "black", hjust = 0, face = "bold"))+
labs(subtitle = "")+
  geom_textbox(aes(x = 3, y = 220, label = countlabel), hjust = 0, vjust = 1,stat = "unique", fill = "white", box.colour = "white")
count
```

dates of establishment streamgraph
```{r}
#count number of new parks per year
numperyear<-alldfs %>% group_by(Established,regionclean) %>% count()

streamlabel1<-"Park establishment picked up pace in the 1950s"
streamlabel2<-"126 parks were established between 1985 and 1989"

streamyear<-ggplot(numperyear, (aes(x = Established, y = n, fill = regionclean)))+
 geom_segment(aes(x = 1985,xend = 1989,y = -30,yend = -30), col = "black")+
  geom_stream(col = "black")+
geom_textbox(aes(x = 1950, y = -10, label = streamlabel1), stat = "unique", hjust = 1, width = unit(0.25, "npc"), fill = "white", box.colour = "white")+
geom_textbox(aes(x = 1980, y = -25, label = streamlabel2), hjust = 1, stat = "unique", fill = "white", box.colour = "white")+
     theme_void()+
    xlab("Date of Park Establishment")+
  theme(legend.position = "none",
        axis.text.x = element_text(color = "black"),
        axis.line.x = element_line(color = "black"),
        axis.title.x = element_text(color = "black", hjust = 0, face = "bold"))

streamyear
```


combine all
```{r}
des = "
1122
1133
"  

ontariomap+count+streamyear+
plot_layout(guides = "keep", design = des)+
plot_annotation(title = "Ontario Provincial Parks",
  caption = "Data: Wikipedia. Chart: @jeff_zeyl",
theme = theme(plot.title = element_text(hjust = 0.5, size = 20),
              legend.position = "right"))
```